Effects of resonant and random excitations on the proton beam in the Large Hadron Collider with applications to active halo control with pulsed hollow electron lenses

### 2018-09-16 resex.tex
Continued with corrections from Referee A: colliders, perveance, MOF/MOD, thus be.

\begin{table*}
  \caption{ Stored beam energy for different past, present and future
    colliders. Each new machine represents a leap in stored beam
    energy.}
  \label{tab:stored_energy}
  \begin{ruledtabular}
    \begin{tabular}{lccccc}
      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016
                                                     \cite{Wenninger:IPAC:2017}
      & LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{Benedikt:JKPS:2016, Benedikt:IPAC:2018} \\
      \colrule
      Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
      Number of bunches & 36 & 2220 & 2808 & 2748 & 10600 \\
      Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
      Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8480 \\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

\begin{table}
  \caption{HL-LHC hollow electron lens parameters, as defined in
    Ref.~\cite{hel_cdr}.}
  \label{tab:hel_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Geometry & Value& Unit\\
      \colrule
      Length, $L$    &  3 & m\\
      % 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
      Desired range of scraping positions & 3.5--9.5 &$\sigma_p$\\
      \colrule
      Magnetic fields & & \\
      \colrule
      Gun solenoid, $B_g$ & 0.2--0.4 & T\\
      Main solenoid, $B_m$ & 2--6 & T\\
      Collector solenoid, $B_c$ & 0.2--0.4 & T\\
      Compression factor, $k \equiv \sqrt{B_m/B_g}$ & 2.2--5.5 &  \\
      \colrule
      Electron gun & & \\
      \colrule
      Peak yield $I_e$ at 10~keV & 5.0 & A\\
      Gun perveance, $P$ & 5 & $\mu$perv\\
      Inner/outer cathode radii, $R_1/R_2$ & 6.75/12.7 & mm\\
      \colrule
      High-voltage modulator & & \\
      \colrule
      Cathode-anode voltage & 10.0 & kV\\
      Rise time (10\%-90\%) & 200 & ns \\
      Repetition rate & 35 & kHz
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

\begin{table*}
  \caption{Beam parameters and machine configuration for the two
    resonant excitation experiments of~2016
    and~2017~\cite{resexmd2016, resexmd2017}. The plane of the
    excitation is abbreviated as H for horizontal, V for vertical and
    H+V for horizontal and vertical at the same time.}
  \label{tab:md_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Parameter & Experiment 2016 & Experiment 2017  \\
      \colrule
      beam &\multicolumn{2}{c}{Beam~1} \\
      beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
      single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
      normalized emittance &\multicolumn{2}{c}{$2.5$--\q{3.5}{\mu m}} \\
      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
      number of bunches & $12 \times 4 = 48$ & $3 \times 72 = 216$ \\
                &  & (+ 1 pilot + 12 nominal) \\\hline
      injection optics, $\beta^* = \q{11}{m}$ & standard optics 2016 & standard optics 2017\\
      Landau-damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}
                                  = \q{+19.6}{A}$ for MOF circuit and}\\
                & \multicolumn{2}{c}{$I_{\mathrm{MO}} = \q{-19.6}{A}$
                  for MOD circuit (standard 2016 settings)} \\ \hline
      working point $(Q_x, Q_y)$ & (64.28, 59.31) & (62.27, 60.295) \\
      chromaticity $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)}\\ \hline
      pulsing patterns  & 7th~turn H & 7th~turn H, V, H+V \\
                & 10th~turn V & 8th~turn H, V, H+V \\
                & &  random  H, V, H+V\\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

The effects of 7th- and \tenthtp\ are also observed in the absence of
machine lattice errors. In this case, the only sources of nonlinearity
are sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer},
suggesting that these nonlinearities are responsible for the
pronounced beam sensitivity. The driven resonances are revealed by the
frequency-map analysis, shown in Fig.~\ref{fig:patternfma}. The
$7 Q_x$ resonance is excited by the \seventhtp, and the $10 Q_x$ and
$10 Q_y$ resonances by the \tenthtp. As octupoles can only drive even
resonances, the sources of the $7 Q_x$ resonances are the sextupoles,
while the octupoles generate the tune footprint. The other pulsing
patterns do not exhibit any increase in losses or emittance growth
without magnetic errors. Their effect can be thus attributed to an
interaction of the excitation with the machine errors, implying also a
sensitivity to the chosen random seed in simulations and to the
specific error distribution in experiments.

>>>>

\begin{table*}
  \caption{Stored beam energy for a few examples of past, present and future
    colliders. New machines represents a leap in stored beam
    energy.}
  \label{tab:stored_energy}
  \begin{ruledtabular}
    \begin{tabular}{lccccc}
      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016
                                                     \cite{Wenninger:IPAC:2017}
      & LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{Benedikt:JKPS:2016, Benedikt:IPAC:2018} \\
      \colrule
      Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
      Number of bunches & 36 & 2220 & 2808 & 2748 & 10600 \\
      Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
      Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8480 \\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

\begin{table}
  \caption{HL-LHC hollow electron lens parameters, as defined in
    Ref.~\cite{hel_cdr}.}
  \label{tab:hel_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Geometry & Value& Unit\\
      \colrule
      Length, $L$    &  3 & m\\
      % 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
      Desired range of scraping positions & 3.5--9.5 &$\sigma_p$\\
      \colrule
      Magnetic fields & & \\
      \colrule
      Gun solenoid, $B_g$ & 0.2--0.4 & T\\
      Main solenoid, $B_m$ & 2--6 & T\\
      Collector solenoid, $B_c$ & 0.2--0.4 & T\\
      Compression factor, $k \equiv \sqrt{B_m/B_g}$ & 2.2--5.5 &  \\
      \colrule
      Electron gun & & \\
      \colrule
      Peak yield $I_e$ at 10~keV & 5.0 & A\\
      Gun perveance, $P$ & 5 & $\mathrm{\mu A / V^{3/2}}$\\
      Inner/outer cathode radii, $R_1/R_2$ & 6.75/12.7 & mm\\
      \colrule
      High-voltage modulator & & \\
      \colrule
      Cathode-anode voltage & 10.0 & kV\\
      Rise time (10\%-90\%) & 200 & ns \\
      Repetition rate & 35 & kHz
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

\begin{table*}
  \caption{Beam parameters and machine configuration for the two
    resonant excitation experiments of~2016
    and~2017~\cite{resexmd2016, resexmd2017}. The plane of the
    excitation is abbreviated as H for horizontal, V for vertical and
    H+V for horizontal and vertical at the same time. MOF and MOD
    refer to the focusing and defocusing Landau octupoles.}
  \label{tab:md_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Parameter & Experiment 2016 & Experiment 2017  \\
      \colrule
      beam &\multicolumn{2}{c}{Beam~1} \\
      beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
      single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
      normalized emittance &\multicolumn{2}{c}{$2.5$--\q{3.5}{\mu m}} \\
      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
      number of bunches & $12 \times 4 = 48$ & $3 \times 72 = 216$ \\
                &  & (+ 1 pilot + 12 nominal) \\\hline
      injection optics, $\beta^* = \q{11}{m}$ & standard optics 2016 & standard optics 2017\\
      Landau-damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}
                                  = \q{+19.6}{A}$ for MOF circuit and}\\
                & \multicolumn{2}{c}{$I_{\mathrm{MO}} = \q{-19.6}{A}$
                  for MOD circuit (standard 2016 settings)} \\ \hline
      working point $(Q_x, Q_y)$ & (64.28, 59.31) & (62.27, 60.295) \\
      chromaticity $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)}\\ \hline
      pulsing patterns  & 7th~turn H & 7th~turn H, V, H+V \\
                & 10th~turn V & 8th~turn H, V, H+V \\
                & &  random  H, V, H+V\\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

The effects of 7th- and \tenthtp\ are also observed in the absence of
machine lattice errors. In this case, the only sources of nonlinearity
are sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer},
suggesting that these nonlinearities are responsible for the
pronounced beam sensitivity. The driven resonances are revealed by the
frequency-map analysis, shown in Fig.~\ref{fig:patternfma}. The
$7 Q_x$ resonance is excited by the \seventhtp, and the $10 Q_x$ and
$10 Q_y$ resonances by the \tenthtp. As octupoles can only drive even
resonances, the sources of the $7 Q_x$ resonances are the sextupoles,
while the octupoles generate the tune footprint. The other pulsing
patterns do not exhibit any increase in losses or emittance growth
without magnetic errors. Their effect can thus be attributed to an
interaction of the excitation with the machine errors, implying also a
sensitivity to the chosen random seed in simulations and to the
specific error distribution in experiments.
